The Dynamic-p Technique with lavaan
======================================================

The purpose of this project is to demonstrate how to fit a single-subject structural equation model within the dynamic p-technique framework. As discussed in the primary manuscript, of the recent introductions to and tutorials on the dynamic p-technique, we believe the method outlined in Little’s (2013) text is the most thorough and approachable for non-statistician behavioral researchers. Thus, the approach described herein is based primarily on Little’s text. Longitudinal structural equation models, such as those estimated with the dynamic p-technique, can become very complex.

```{r set-options, echo = FALSE, cache = FALSE}
options(width = 100)
```

First we'll load the data and open our first two packages.

```{r, warning = F, message = F}
library(readxl)
library(tidyverse)

d <- read_excel("Lindsey.xlsx")

glimpse(d)
```

Note how the data are in the long format. That is, although we have one participant, "Lindsey", her data are presented in 103 rows, each corresponding to a different calendar day. 

The variable for calendar days, `Date`, is descriptively useful, but cumbersome for formal statistical analysis. Here we'll want to augment the data slightly by adding a numeric `day` index, ranging from 0 to 102.

```{r}
d <-
  d %>% 
  mutate(day = 1:n()-1) %>% 
  select(day, everything())
```

## Descriptive statistics

If you wanted to get a sense of the distributions of the ASRS items, histograms might be handy.

```{r, fig.width = 8, fig.height = 3.5}
d %>% 
  select(A3:A17) %>% 
  gather(item, rating) %>% 
  
  ggplot(aes(x = rating)) +
  geom_histogram(binwidth = 1, fill = "chartreuse", color = "grey92", size = 1/5) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~item, ncol = 4)
```

Happily, their distributions look reasonable. You can get a sense of their values over time with a sequentially-color-coded tile plot.

```{r, fig.width = 10, fig.height = 2.5, warning = F, message = F}
library(viridis)

d %>% 
  select(Date, A3:A17) %>% 
  gather(item, rating, -Date) %>%
  mutate(rating = factor(rating, levels = c(1:5)),
         item = factor(item, levels = c("A3", "A8", "A10", "A13", "A14", "A16", "A17"))) %>% 
  
  ggplot(aes(x = Date, y = item)) +
  geom_raster(aes(fill = rating)) +
  scale_fill_viridis(NULL, discrete = TRUE, option = "D",
                     guide = guide_legend(direction = "horizontal",
                                          nrow = 1)) +
  scale_x_datetime(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = NULL, y = NULL) +
  theme(panel.grid = element_blank(),
        axis.text.y = element_text(hjust = 0),
        axis.ticks.y = element_blank(),
        legend.position = "top")
```

And with the [psych package](https://cran.r-project.org/web/packages/psych/index.html), we can use the `describe()` function to get the typical descriptive statistics.

```{r, warning = F, message = F}
library(psych)

d %>% 
  select(A3:A17) %>% 
  describe()
```  

Recall that the items in the ASRS correspond to the 18 criterion A symptoms, with the first nine classified as Inattentive and the last nine as hyperactive/impulsive. For the sake of this example, we'll focus on Lindsey's hyperactive/impulsive symptoms, ASRS items 10, 13, 14, 16, and 17. The following code drops the other two, items 3 and 8.

```{r}
d <-
  d %>% 
  select(everything(), -A3, -A8)
```

## P-technique CFA

Here we open our primary statistical package, [lavaan](https://cran.r-project.org/web/packages/lavaan/index.html), which you might learn more about [here](https://dornsife.usc.edu/assets/sites/210/docs/GC3/lavaan_tutorial.pdf) or [here](http://lavaan.ugent.be/tutorial/index.html).

```{r, message = F, warning = F}
library(lavaan)
```

### 0-lag.

The simplest starting place is a p-technique factor analysis. Using this approach, the model syntax looks much like that of a typical group-level CFA. 

```{r}
CFA_0_lag <- '
H =~ NA*A10 + A13 + A14 + A16 + A17

# Standardize the variance
H ~~ 1*H
'
```

Here we fit the model.

```{r}
fit.CFA_0_lag <- 
  cfa(CFA_0_lag, 
      data = d,
      estimator = "MLR",
      missing = "ML")
```

The summary, including the typical fit indices and 95% CIs for good measure:

```{r}
summary(fit.CFA_0_lag, 
        fit.measures = T,
        ci = T)
```

### 1-lag.

Before we proceed to fit a dynamic-p CFA, we'll need to lag our data file. First, we'll add a row to the end of the data. Because this row corresponds to `Date = "2016-05-11"`, a day for which we don't have any information other than it was a Wednesday, we'll insert NAs into most of the columns.

```{r}
tail(d)

d_lagged <-
  d %>% 
  add_row(., day = 103, Date = "2016-05-11 00:00:00", 
          Mon = 0, Tues = 0, Wed = 1, Thur = 0, Fri = 0, Sat = 0, Sun = 0, 
          Meds = NA, 
          A10 = NA, A13 = NA, A14 = NA, A16 = NA, A17 = NA)
```

With the `lag()` function, we'll add lagged values for our `Meds` dummy and our ASRS items. For each of the lagged columns, we'll add the suffix `_1` to help differentiate them from the original columns.

```{r}
(
  d_lagged <-
  d_lagged %>% 
  mutate(Meds_1 = lag(Meds),
         A10_1  = lag(A10),
         A13_1  = lag(A13),
         A14_1  = lag(A14),
         A16_1  = lag(A16),
         A17_1  = lag(A17))
  )
```

Now we have our lagged data, `d_lagged`, we are ready to specify and fit the 1-lag CFA model.

Because of the way the data are copied to create a lagged dataset, model estimation involving multiple lags entails a number of specific constraints because the information across the lags is essentially equivalent (Little, 2013). The first few of those constraints concern factorial invariance. Typical group-level longitudinal CFAs require that analysts assess the extent to which the factors are invariant across time (for detailed discussions, see Brown, 2015; Little, 2013; Widaman, Ferrer, & Conger, 2010). Factorial invariance across time suggests the variables of interest were reliably measured across time and that the constructs themselves were stable. For dynamic p-technique models, we expect "strict factorial invariance" (Little, 2013), which entails that the item loadings, item intercepts, and residual variances are equivalent across lags (herein termed constraints a, b, and c, respectively). As is typical of longitudinal CFAs, residual covariances of the same items measured across lags are estimated and constrained to equality when separated by the same number of lags. Though not applicable to the models herein, we might call this constraint d. 

The above four constraints all concern CFA measurement models (i.e., parameters used to identify the latent variables). The dynamic p-technique also entails that parameters in the structural model (i.e., parameters among latent variables and, when applicable, observed covariates) are equivalent across lags. Specifically, latent variable means for a given construct might be constrained to equality across lags (constraint e). Regressions between variables are also constrained to equality when separated by the same number of lags (constraint f). Due to the nature of the structural model in these examples, we will not impose these constraints. But see Little's text for details.

For researchers unaccustomed with equality constraints such as these, we recommend first estimating an unconstrained model and then specifying the various equality constraints one at a time. Doing so may help clarify the implications of each constraint.

Anyway, here's the model:

```{r}
CFA_1_lag <- '
# Loadings
# The `l[i]*` portions constitute constraints a

H0 =~ l1*A10 +   l2*A13 +   l3*A14 +   l4*A16 +   l5*A17
H1 =~ l1*A10_1 + l2*A13_1 + l3*A14_1 + l4*A16_1 + l5*A17_1

# LV variances
H0 ~~ 1*H0
H1 ~~ NA*H1 # Because of the structural model, the resitual variance for H1 is freely estimated

# item intercepts
# The `i[i]*` portions constitute constraints b

A10 ~ i1*A10
A13 ~ i2*A13
A14 ~ i3*A14
A16 ~ i4*A16
A17 ~ i5*A17

A10_1 ~ i1*A10_1
A13_1 ~ i2*A13_1
A14_1 ~ i3*A14_1
A16_1 ~ i4*A16_1
A17_1 ~ i5*A17_1

# residual variances
# The `rv[i]*` portions constitute constraints c

A10 ~~ rv1*A10
A13 ~~ rv2*A13
A14 ~~ rv3*A14
A16 ~~ rv4*A16
A17 ~~ rv5*A17

A10_1 ~~ rv1*A10_1
A13_1 ~~ rv2*A13_1
A14_1 ~~ rv3*A14_1
A16_1 ~~ rv4*A16_1
A17_1 ~~ rv5*A17_1

# cross-lag residual covariances
A10 ~~ A10_1
A13 ~~ A13_1
A14 ~~ A14_1
A16 ~~ A16_1
A17 ~~ A17_1

# Structural model
H1 ~~ H0
'
```

We fit it, here.

```{r}
fit.CFA_1_lag <- 
  cfa(CFA_1_lag, 
      data = d_lagged,
      estimator = "MLR",
      missing = "ML",
      std.lv = T)
```

The summary:

```{r}
summary(fit.CFA_1_lag, 
        fit.measures = T,
        ci = T)
```

The model fits the data great.

## Dynamic-p SEM

Finally, we can use `Meds_1` to predict lag-1 ADHD values, `H1`, while still controlling for the previous day's ADHD values (i.e., `H1 ~ H0`). The same measurement invariance constraints are imposed.

```{r}
SEM_1_lag <- '
# Loadings
H0 =~ l1*A10 +   l2*A13 +   l3*A14 +   l4*A16 +   l5*A17
H1 =~ l1*A10_1 + l2*A13_1 + l3*A14_1 + l4*A16_1 + l5*A17_1

# LV variances
H0 ~~ 1*H0
H1 ~~ NA*H1 # Because of the structural model, the resitual variance for H1 is freely estimated

# item intercepts
A10 ~ i1*A10
A13 ~ i2*A13
A14 ~ i3*A14
A16 ~ i4*A16
A17 ~ i5*A17

A10_1 ~ i1*A10_1
A13_1 ~ i2*A13_1
A14_1 ~ i3*A14_1
A16_1 ~ i4*A16_1
A17_1 ~ i5*A17_1

# residual variances
A10 ~~ rv1*A10
A13 ~~ rv2*A13
A14 ~~ rv3*A14
A16 ~~ rv4*A16
A17 ~~ rv5*A17

A10_1 ~~ rv1*A10_1
A13_1 ~~ rv2*A13_1
A14_1 ~~ rv3*A14_1
A16_1 ~~ rv4*A16_1
A17_1 ~~ rv5*A17_1

# cross-lag residual covariances
A10 ~~ A10_1
A13 ~~ A13_1
A14 ~~ A14_1
A16 ~~ A16_1
A17 ~~ A17_1

# Structural model
H1 ~ H0 + Meds_1
'
```

Fitting the model

```{r}
fit.SEM_1_lag <- 
  cfa(SEM_1_lag, 
      data = d_lagged,
      estimator = "MLR",
      missing = "ML",
      std.lv = T)
```

The summary:

```{r}
summary(fit.SEM_1_lag, 
        fit.measures = T,
        ci = T)
```

Recall that the latent variables are in a standardized metric. Because `Meds_1` is a dummy variable, this puts it's coefficient in a Cohen's $d$ like metric.

## References

Brown, T. A. (2015). *Confirmatory factor analysis for applied research*. New York, NY: Guilford Press.

Little, T. D. (2013). *Longitudinal structural equation modeling*. New York, NY: The Guilford Press. 

Widaman, K. F., Ferrer, E., & Conger, R. D. (2010). Factorial invariance within longitudinal structural equation models: Measuring the same construct across time. *Child Development Perspectives, 4*, 10-18. doi:10.1111/j.1750-8606.2009.00110.x
